### 2021-06-29 paper/header.tex
Add files for paper

% **************GENERATED FILE, DO NOT EDIT**************

\title{DeconvOptim.jl - Image Deconvolution with Julia}

\author[1, 2]{Felix Wechsler}
\author[1, 2, 3]{Rainer Heintzmann}
\affil[1]{Faculty of Physics and Astronomy, Friedrich-Schiller-University,Jena, Germany}
\affil[2]{Leibniz Institute of Photonic Technology, Albert-Einstein-Straße 9,07745 Jena, Germany}
\affil[3]{Institute of Physical Chemistry and Abbe Center of Photonics, Friedrich-Schiller-University, Helmholtzweg 4, Jena, Germany}

\keywords{Julia, Image Processing, Deconvolution, Microscopy}

\hypersetup{
pdftitle = {DeconvOptim.jl - Image Deconvolution with Julia},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Felix Wechsler, Rainer Heintzmann},
pdfkeywords = {Julia, Image Processing, Deconvolution, Microscopy},
}