### 2021-06-30 paper/paper.tex
Update paper content

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{siunitx}
\usepackage{here}
\newcommand\todo[1]{\textcolor{red}{\textbf{TODO: #1}}}
\begin{document}

\input{header}

\maketitle

\begin{abstract}

Deconvolution is a versatile method to enhance the quality of signals measured with systems which can 
be mathematically expressed as a convolution of a system's response function with a signal. 
In this paper, we present \verb|DeconvOptim.jl|, a flexible toolbox written in Julia to deconvolve multi-dimensional signals which have been
degraded by a multi-dimensional signal response function. 
\verb|DeconvOptim.jl| works both on CPUs and GPUs and utilizes recent advancements in Julias Automatic Differentiation ecosystem. The performance on CPUs surpasses the one of existing open source libraries. 
The library is applicable to one dimensional time series datasets but also to multi-dimensional microscopy imaging datasets.
\end{abstract}

\todo{fix code formatting}

\section{Introduction}
Deconvolution has been a long addressed problem especially in astronomical imaging, digital signal processing or microscopy imaging.
The assumption is that images are degraded by some kind of blur which can be described as a convolution between a kernel $h$ and the image $S$:
\begin{equation}
    I(\mathbf r) = (h * S)(\mathbf r)
\end{equation}
$h$ is often called the point spread function (PSF) since it characterizes how the systems maps a point in the object to a blurred area
in the image.
Often this convolution operation is disturbed by different types of noise where the most common are additive Gaussian noise due to the read out of the camera sensors or Poisson shot noise due to the quantum nature of photons:
\begin{align}
    Y_{\text{Poisson}}(\mathbf r) &= \text{Poisson}[(h * S)(\mathbf r)]\\
    Y_{\text{Gauss}}(\mathbf r) &= (h * S)(\mathbf r) + \text{Gauss}_{\mu, \sigma}(\mathbf r)
\end{align}
Due to the noise and the low pass filter effect of the PSF, deconvolution is an ill-posed problem which cannot be solved directly.\\
A well known approach, because of its simplicity, has been the iterative Richardson-Lucy deconvolution algorithm \cite{Richardson:72}, \cite{lucy:74} for Poisson noise degraded signals.
Another common method is the non-iterative Wiener filter \cite{wiener2013extrapolation} because of its predictable runtime.
Wiener filter assumes additive noise which is often not true, especially in low light conditions as in microscopy or astronomy. 
Recently machine learning based methods have become more and more favoured in deconvolution methods because
of the possibility to add learned properties of similar images to the algorithm \cite{Kruse_2017_ICCV}.
However, those implementations are usually restricted to cases where experimental datasets are available for the characterization and training.
\verb|DeconvOptim.jl| offers several deconvolution algorithms like the iterative Richardson-Lucy algorithms but also
a much more flexible method which is based on a loss function being minimized by a gradient descent.
Also this toolbox is not restricted in dimensionality and therefore can be applied to deconvolution problems in many different fields.

\section{Efficient Convolutions}
The mathematical definition of a convolution is
\begin{equation}
    (h * S)(\mathbf r) = \int_{-\infty}^{\infty} h(\mathbf r - \mathbf x) \cdot S(\mathbf x) \, \mathrm d \mathbf x.
    \label{eq:conv}
\end{equation}
To obtain efficient deconvolution algorithms we also need a efficient way to calculate the convolution operations.
In these times, convolution operations of spatially small kernels $h$ can be calculated on GPUs very efficiently.
However, often the blurring kernel $h$ is not spatially small and might be also multi-dimensional. In that case, FFT based convolutions
outperform sliding kernels. In microscopy and astronomy that is often the case, since a defocus broadens the PSF.

\subsection{FFT Based Convolution}
    The basis for fast convolution operations lies in the convolution theorem. It states that
    \begin{equation}
        \mathcal F\left\{(h * S)\right\}(\mathbf k) = \mathcal F\{h\} (\mathbf k) \cdot \mathcal F\{S\} (\mathbf k),
        \label{eq:fftconv}
    \end{equation} 
        which means that a Fourier transform of a  convolution operation is equivalent to a point wise multiplication of the Fourier transforms 
        of $h$ and $S$.
        Fortunately, the Fourier transform operations can be calculated even for moderate large 3D arrays quickly on modern computing machines via
        the Fast Fourier Transform (FFT).
        In Julia, \verb|FFTW.jl| offers a convenient interface to the FFTW library \cite{FFTW05}.
        The time complexity to calculate \autoref{eq:fftconv} is $\mathcal{O}(N \cdot \log N)$ for a one dimensional discretized dataset with $N$ samples.
        For a 2D dataset with size $N \times M$ it is $\mathcal{O}(N \cdot M \log(N + M))$ and similar for higher dimensions.

\subsection{Sliding Convolution Windows}
Convolutions can be also calculated with a sliding kernel in real space which is the direct approach to
solve the integral in \autoref{eq:conv}.
That mechanism is the working principle
behind the convolutions in neural nets. The reason is that this operation can be parallelized easily and therefore 
is suitable on GPUs. However, the computational complexity for a 2D kernel with size $K \times L$ applied on an image with
$N \times M$ data points is $\mathcal O(N \cdot M \cdot K \cdot L)$. It is clear that for larger kernel sizes ($K,L \gg 1$) the sliding method 
has worse computational complexity than the FFT based convolutions. The many cores of a GPU can mitigate that to a certain extent
but since the FFT (CuFFT in the CUDA library \cite{CUDA}) can be also executed in parallel fashion, both approaches profit from parallelization.

\subsection{Wrap-Around Artifacts in FFT based Convolutions}
    One drawback of the FFT based convolution are wrap-around artifacts. The reason is, that the FFT operation calculates the discrete
    Fourier transform (DFT) which inherently assumes periodic data. If one applies a FFT on a finite data set the assumption is that
    this finite data set repeats itself. 
    To weaken the wrap-around artifacts in the convolution one can zero or mean pad the image before convolution and remove the boundaries afterwards.

    \begin{figure}[h]
        \centering
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_no_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_no_pad.png}};
                \draw[color=red] (-0.91, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
            \end{tikzpicture}
            \caption{Without padding}
            \label{fig:withoutpad}
        \end{subfigure}%
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_pad.png}};
                \draw[color=red] (-0.90, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
            \end{tikzpicture}
            \caption{With padding.}
            \label{fig:withpad}
        \end{subfigure}
        \caption{Comparison between two blurred images. The image a) is a straightforward FFT based blurred image.
        The zoomed region shows dark artifacts at the left side originated from the right hand side. Also at the top of the image some dark regions from
        the bottom scattered into the bright sky.
        Image b) was padded before blurring with 10 pixels on each side with value 0.7. The blur kernel was a Gaussian kernel
        with standard deviation $\sigma=\SI{3}{pixel}$.}
        \label{fig:}
    \end{figure}

\subsection{Preferred Convolution}
The current state of \verb|DeconvOptim.jl| uses the FFT approach to calculate the convolution.
However, in principle one could add the possibility to change to spatially based convolutions.
But since the runtime gets especially long in multi-dimensional signal deconvolution where FFT based methods
are faster, we always use FFT based convolutions.

\section{Inverse Modelling}
In \autoref{fig:pipeline} we show our approach to deconvolution.
The general idea is to create a (physical) forward model describing the measurement process.
\begin{figure}[h]
    \centering
    \includegraphics[width = .5\textwidth]{figures/pipeline.pdf}
    \caption{}
    \label{fig:pipeline}
\end{figure}

At the beginning of the optimization we start with an initial guess $x(\mathbf r)$. 
Via a mapping function we can impose some restrictions to the final solution (e.g. non-negativity, certain value intervals).
The forward model is then applied to the mapped reconstruction. We provide a default function where 
the forward function is a convolution but we also allow to plug in a different forward model into the pipeline.
This output of the forward function is compared
with our noisy measurement under a certain loss function.
Since deconvolution is an ill-posed problem we regularize the loss function with a regularizer.
Afterwards we calculate the gradient of the loss function with respect to reconstruction $x(\mathbf r)$ via automatic differentation (AD).
The values of the total loss and the gradient are then used by an optimizer (to minimize the total loss) to apply some kind of gradient descent optimization to get a proper reconstruction.\\

As typical for Julia packages, except \verb|FFTW.jl| and \verb|CUDA.jl| \cite{besard2018juliagpu}, all code is written in Julia.
Therefore it is easy to replace the forward model with any Julia code, as long as we can 
automatically differentiate through the code. The gradient step is currently calculated with the reverse mode AD of \verb|Zygote.jl| \cite{Zygote.jl-2018}. 
As default for the optimization we use the L-BFGS \cite{LBFGS} routine provided by \verb|Optim.jl| \cite{mogensen2018optim}.
L-BFGS is suitable for optimization with many million parameters as it only stores a sparse representation of the inverse Hessian matrix
which is built from a finite number of past function and gradient evaluations. 
In comparison to the Richardson-Lucy algorithm results are often good after 10 to 30 iterations whereas Richardson-Lucy takes
a few hundreds iterations. 
However, as we will see later due to the more complex optimization routine the speed-up in using L-BFGS is for simple 
deconvolution minor.

\subsection{Mapping Function}
    Mapping functions are an easy way to constrain the resulting reconstruction to certain values.
    For example, if the optimizer optimizes the variables $\text{rec}$ which are then squared, we know that
    $\text{rec}^2$ is definitely non-negative. Instead of directly using $\text{rec}$ we compare
    $\text{rec}^2$ with the measurement $Y$. This mapping function does not change the outcome as long
    as it is an surjective function for the desired number space (e.g. a parabola is non-negative and covers all positive values) 
    and does not change the convexity of a problem.
    Functions like shifted $\tanh$ are possible as well since it is an bijective function in $[-1, 1]$ and monotonic increasing
    in the whole definition space.
    Hence the final result is $\text{rec}^2$ in case of the quadratic mapping.

\subsection{Wrap-around Artifacts Suppression}
    To prevent that wrap-around artifacts occur during deconvolution we can instead
    of reconstructing an object having the same size as the measurement, reconstruct
    an object having a slightly larger size. This is shown in \autoref{fig:wrap_around}.
    \begin{figure}[h]
        \centering
        \begin{tikzpicture}
            \draw[fill=red!20!white] (0,0) rectangle (4,4);
            \node[color=red] at (2, 3.8) {Padded area};
            \draw[fill=gray!40!white] (0.5,0.5) rectangle (3.5,3.5); 
            \node[text width=2.5cm, color=black] at (2, 2) {area compared with measurement};
        \end{tikzpicture}
        \caption{Reconstructed data is padded before deconvolution but only the red region is compared with the measurement.}
        \label{fig:wrap_around}
    \end{figure}
    Consequently, we only compare the grey part of the reconstruction with the measurement.
    Through the AD and the convolution operations we still optimize for the whole area (including the red one).
    With that padding we prevent that images on the boundary influence in circular manner the reconstruction.

\section{Loss Function}
    Besides the forward model, the loss function is the most critical part in the optimization.
    It has to take account for the noise created by the measurement process or device.

\subsection{Poisson Data Term}
    As already mentioned, Poisson shot noise is often the dominant source of noise hence we 
    explain the details.
    The key idea is to interpret the measurement as a stochastic process. Our aim is to find a deconvolved image which describes as accurate as possible the measured image.
    Mathematically the probability for a certain measurement $Y$ under Poisson noise is

\begin{equation}
    p(Y(\mathbf r)|\mu(\mathbf r)) = \prod_r \frac{\mu(\mathbf r)^{Y(\mathbf r)}}{\Gamma(Y(\mathbf r) + 1)} \exp(- \mu(\mathbf r))
\end{equation}

where $Y$ is the measurement, $\mu$ is the expected measurement (ideal measurement without noise) and $\Gamma$ is the generalized factorial function.
In the deconvolution process we get $Y$ as measured input and want to find the ideal specimen $S$ which results in a measurement $\mu(\mathbf r) = (S * \text{PSF})(\mathbf r))$.
Since we want to obtain for the best reconstruction, we want to find a $\mu(\mathbf r)$ so that $p(Y(\mathbf r) | \mu(\mathbf r))$ gets as large as possible. Because that means
that we find the specimen which describes the measurement with the highest probability.
Instead of maximizing $p(Y(\mathbf r) | \mu(\mathbf r))$ a common trick is to minimize $- \log(p(Y(\mathbf r)|\mu(\mathbf r)))$. 
Mathematically, the optimization of both functions provides same results but the latter is numerically more stable.

\begin{align}
&\underset{S(\mathbf r)}{\arg \min} (- \log(p(Y(\mathbf r)|\mu(\mathbf r)))) =\\ &\underset{S(\mathbf r)}{\arg \min} \sum_r (\mu(\mathbf r) + \log(\Gamma(Y(\mathbf r) + 1)) - Y(\mathbf r) \log(\mu(\mathbf r))
\end{align}

which is equivalent to

\begin{equation}
    \underset{S(\mathbf r)}{\arg \min}\, L = \underset{S(\mathbf r)}{\arg \min} \sum_r (\mu(\mathbf r)  - Y(\mathbf r) \log(\mu(\mathbf r))
    \label{eq:poisson_loss}
\end{equation}

since the second term only depends on $Y(\mathbf r)$ but not on $\mu(\mathbf r)$.
The gradient of $L$ with respect to $\mu(\mathbf r)$ is simply

\begin{equation}
    \nabla L = 1 - \frac{Y(\mathbf r)}{\mu(\mathbf r)}
    \label{eq:gradpoisson}
\end{equation}

The function $L$ and the gradient $\nabla L$ are needed for any gradient descent optimization algorithm.
The numerical evaluation of the Poisson loss can lead to issues. Since $\mu(r)=0$ can happen for a measurement with zero intensity background. However, the loss is not defined for $\mu \leq 0$. In our source code we set all intensity values below a certain threshold $\epsilon$ to $\epsilon$ itself. This prevents the evaluation of the logarithm at undefined values.
The final source code has a modified version of \autoref{eq:gradpoisson} as registered gradient for \autoref{eq:poisson_loss}.
The reason is, that for values of $\mu(\mathbf r) \leq 0$ the equation is not defined and has to be handled separately.

\subsection{Regularizer}
    The regularizer is an additional part in the total loss to add some additional priors like sparsity and smoothness.
    Most real samples are smooth over large regions and such regularizers enhance the quality of the reconstruction.
    In our toolbox we currently implement several regularizer like (smoothed) Total variation, Good's roughness \cite{Verveer:98}, \cite{Good:71} and Tikhonov. 
    \subsubsection{Total Variation as example}
        As the name suggests, Total variation tries to penalize variation in the image intensity. Therefore it sums up the gradient strength at each point
        of the image:
        \begin{equation}
            \text{Reg}(S(\mathbf r)) = \sum_r  |(\nabla S)(\mathbf r)|
        \end{equation}
        Since we look at the magnitude of the gradient strength, this regularizer is anisotropic.
        In 2D TV is defined like: 
        \begin{equation}
        \begin{split}
            \text{Reg}(S(\mathbf r)) = \sum_{x,y} \big[& |S(x + 1, y) - S(x, y)|^2 \\ 
                                                & + |S(x, y + 1) - S(x, y)|^2\big]^{\frac12}
        \end{split}
        \end{equation}
        In many frameworks such a regularizer is implemented via functions like \lstinline{circshift}. 
        However, such approaches create often several copies of the data. 
        We currently use \verb|Tullio.jl|\footnote{\url{https://github.com/mcabbott/Tullio.jl}} to create efficient functions 
        calculating the regularizer and the gradient. For simple expressions
        such as below, \verb|Tullio.jl| registers an analytical gradient and therefore allows to apply the same efficient mechanisms 
        to the gradient calculation which have been used for the regularizer evaluation itself \todo{ask McAbott for reason}.
        A (smoothed) Total variation with \verb|Tullio.jl| is defined here:
        \begin{lstlisting}[language = Julia]
function total_variation(arr, ϵ=eltype(arr)(1e-8))
    @tullio r = sqrt(ϵ + 
                abs2(arr[i,j] - arr[i+1,j]) +
                abs2(arr[i,j] - arr[i,j+1]))
end
        \end{lstlisting}
        \lstinline{ϵ} is needed for regions where \lstinline{arr == 0} because otherwise the gradient return infinite values.
        Note, that definition here just demonstrates how the regularizer can be written down.
        In our toolbox we allow for much more options. Since \verb|Tullio.jl| cannot handle generic multi-dimensional data yet
        we create regularizers which can vary in step size, dimensionality and
        different weighting in dimensions via Julia's metaprogramming capability.
        A call like 
        \begin{lstlisting}[language = Julia]
julia> using DeconvOptim

julia> reg = TV(num_dims=4, sum_dims=[1,2,3],
                weights=[1,1,2])
#169 (generic function with 1 method)

julia> reg(randn((10,20,30,40)))
417810.11332404887
        \end{lstlisting}
        results in a regularizer accepting four dimensional arrays. However, the regularization only happens along the first three dimension
        and the third dimensions is weighted differently.
        Such options are available for most of the regularizers implemented in our toolbox.

\section{Experiments}
    In this section we show some experiments and compare parts of the software to other solutions.
    All experiments were executed on Ubuntu 20.04 using a AMD Ryzen 5 5600X 6-Core processor (12 threads) with \SI{32}{GB} DDR4 RAM and Julia 1.6.1.
    If possible, all threads were used.
    As GPU accelerator we used a GeForce RTX 2060 SUPER with \SI{8}{GB} memory and CUDA 11.3 and CUDA.jl 3.3.1.

    \subsection{One-Dimensional Time Series}
        First, we want to demonstrate the deconvolution of an artificial one dimensional dataset blurred with a certain PSF.
        Additionally, we want to demonstrate the mechanism to reduce wrap-around artifacts in the deconvolution.
        On the left hand side in \autoref{fig:1D_dataset} we can see a data series which has been simulated 
        as a convolution with a PSF and degradation with additive Gaussian noise.
        \begin{figure}[H]
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/1d_plot_data_blurred.pdf}
                \caption{}
                \label{fig:1D_dataseta}
            \end{subfigure}%
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/1d_plot_data_res_pad_vs_nopad.pdf}
                \caption{}
                \label{fig:1D_datasetb}
            \end{subfigure}
            \caption{1D data which has been convolved with a PSF and was degraded with additive Gaussian noise with $\sigma=1.0$.}
            \label{fig:1D_dataset}
        \end{figure}
        On the right hand we can see the results of the following deconvolution:
        \begin{lstlisting}[language=Julia]
r_p, o = deconvolution(measured, psf, loss=Gauss(),
     iterations=12, regularizer=TV(num_dims=1),
     λ=0.05, padding=1)
r, o = deconvolution(measured, psf, loss=Gauss(), 
     iterations=12, regularizer=TV(num_dims=1),
     λ=0.05, padding=0)
        \end{lstlisting}
    The first statement applies a total padding of $1$ meaning that the size is doubled and therefore the measured data is padded.
    As we can see, the blue curve in \autoref{fig:1D_datasetb} does not show these intense spikes at the boundaries as the green curve shows.
    Also we notice that the deconvolved data in \autoref{fig:1D_datasetb} does not show all the details which are visible in the ground truth.
    The reason is that the PSF is a low pass filter and therefore only frequencies which have been passed through the filter 
    can be enhanced in terms of contrast in a deconvolved result.

    \subsection{Influence of Regularizer}
        In this part we want to demonstrate the influence of the regularizer.
        \autoref{fig:fabioa} is a simulated imaged which has been affected by Poisson shot noise and blurring by a PSF.
        \autoref{fig:fabiob} shows the results after 15 iterations of the deconvolution. 
        We see that the image is heavily affected by some artifacts which usually occur after too many iterations without a regularizer.
        \autoref{fig:fabioc} shows the results after 15 iterations of the deconvolution but regularized with a Good's roughness regularizer.
        The total image quality is much better. A larger regularizer weight dampens the artifacts even more but also results in lower
        resolution.
        \begin{figure}[h]
            \begin{subfigure}[b]{.166\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/fabio_blurry.png}
                \caption{Measured}
                \label{fig:fabioa}
            \end{subfigure}%
            \begin{subfigure}[b]{.166\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/fabio_deconv_no_reg_iter_30.png}
                \caption{Without regularizer}
                \label{fig:fabiob}
            \end{subfigure}%
            \begin{subfigure}[b]{.166\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/fabio_deconv_regularizer_005_iter_30.png}
                \caption{With regularizer}
                \label{fig:fabioc}
            \end{subfigure}%            
            \caption{a) has been convolved with a PSF and affected by Poisson shot noise (1000 photons expected for the brightest pixel). b) has been deconvolved for 15 iterations without an regularizer whereas c) was regularizer with Good's roughness.}
            \label{fig:image}
        \end{figure}
        In \autoref{fig:fft_data_image} we show the absolute value of the FFT of the data. 
        The meaning of this plot is to judge how the deconvolution enhances the contrast of frequencies which have 
        been blurred by the convolution before.
        The blue curve is the measured data and is below the red ground truth curve. This is due to the low pass filter effect of the PSF.
        The measured curve also contains high frequency information which has been added due to noise and would not be expected 
        after a pure convolution.
        \begin{figure}[h]
            \centering
            \includegraphics[width = .4\textwidth]{figures/frequencies_img.pdf}
            \caption{This plot shows the absolute value of the FFT of the data.}
            \label{fig:fft_data_image}
        \end{figure}
        We see that the orange curve also contains high spatial frequency information. The reason is, that the deconvolution does amplify also noise
        without a regularizer. The green curve shows that the regularizer indeed dampens the high frequencies and results in a visually much
        nicer image which is also closer to the ground truth

    \subsection{Multi-Dimensional Microscopy Data}
        To show the full ability, we deconvolve a four dimensional dataset where the measured volume consists of a 3D datasets
        with three different color channels which is shown in \autoref{fig:result_3d}.
        The PSF is only three dimensional.
        Julia's broadcast mechanism handles that the PSF is only three dimensional and therefore
        we only need to specify over which dimensions the convolution should happen.
        \begin{figure}[h]
            \begin{subfigure}{.25\textwidth}
                \centering
                \includegraphics[width = .95\textwidth]{figures/deconvolved_rgb.png}
            \end{subfigure}%
            \begin{subfigure}{.25\textwidth}
                \centering
                \includegraphics[width = .95\textwidth]{figures/deconvolved_rgb.png}
            \end{subfigure}
            \caption{Slice of a 3D dataset with three different colors. a) is the raw data and b) is the
                    deconvolved image.}
            \label{fig:result_3d}
        \end{figure} 
        Clearly we can see that the deconvolution removed a lot of the background blur.

    \subsection{Performance Comparison}
        Since Julia is often advertised with great performance we compare parts and the full deconvolution to other implementations in
        terms of their total performance.
        \subsubsection{Regularizer}
            In \autoref{tab:reg} we can see the runtimes for different TV regularizer implementations. As we can see, \verb|Tullio.jl| outperforms a plain 
            Julia implementation by more than order of magnitude. Partially, that is caused by the reason that \verb|Tullio.jl| can make 
            use of all threads whereas Julia's built in broadcast mechanism is not multi threaded.
            By \textit{plain Julia} we are referring to an implementation which is written with standard Julia Base functions.
            The same plain Julia implementation was then used together with \verb|CUDA.jl|
            to make use of the GPU capabilities. We see that there is a minor benefit in using a GPU for the regularizer instead of 
            the \verb|Tullio.jl| version on the CPU. Unfortunately, GPU support by \verb|Tullio.jl| is still experimental and especially on the gradient step 
            we could not achieve reasonable results with it.
            \begin{table}[h]
                \begin{tabular}{l l l l l}
                    & plain & $\nabla$ plain & \verb|Tullio.jl| & $\nabla$ \verb|Tullio.jl| \\
                \hline
                CPU & \SI{83.9}{\milli\second} & \SI{7.18}{\second} & \SI{5.53}{\milli\second}  & \SI{119.8}{\milli\second}   \\ 
                GPU & \SI{2.5}{\milli\second}  & \SI{22.9}{\milli\second}& \SI{4.00}{\milli\second}  & \SI{12.53}{\second} \\
                \hline
                \hline
                    & Jax & $\nabla$ JAX && \\
                \hline
                CPU & \SI{48}{\milli\second}   & \SI{364}{\milli\second}&   & \\
                GPU & \SI{887}{\micro\second}   & \SI{10.2}{\milli\second}&   &
                \end{tabular}
                \caption{Runtimes of the Total Variation regularizer for different Julia versions and Jax. 
                    We used a 3 dimensional Float32 array with size $300 \times 300 \times 300$.  }
                \label{tab:reg}
            \end{table}
            We also compare the performance to an implementation written in Python with the Framework JAX \cite{jax2018github}. 
            The GPU version seems to outperform the Julia version by roughly a factor of $2 - 3$ in both forward and gradient pass.
            However, on the CPU \verb|Tullio.jl| is considerably faster. 
            In general one could always squeeze more performance out of the gradient calculations if the kernels would be hand optimized.
            In our package we especially try avoid that because of the multi-dimensionality it is much more convenient
            to rely on AD and as shown above the performance is usually acceptable.

    \subsection{Full Deconvolution}
        The aim of this comparison is not to compare the quality of the results but the computing time to achieve similar reconstructions. The algorithms used in \verb|DeconvOptim.jl| are not fundamentally different to a plain Richardson-Lucy deconvolution and therefore the results will 
        not exceed those results significantly in terms of quality.\\
        In \autoref{tab:deconv} we show the computing time for different algorithms and packages and in \autoref{tab:deconvprop} the basic settings
        in each package.
        One should keep in mind that such performance comparison are usually biased since we are not
        familiar with all options which should be choosen to achieve best performance.
        The comparison should give a rough overview which runtime is expected for each toolbox.\\
        The different packages used were \verb|ThreeDeconv.jl| \cite{ikoma2018convex}, \verb|Huygens| Professional version 21.04 (Scientific Volume Imaging, The Netherlands, \url{http://svi.nl}), \verb|DeconvolutionLab2| \cite{deconvlab2} and \verb|GenericDeconv|.

        \begin{table}[h]
            \begin{tabular}{l l l}
                &  CPU in s& GPU in s\\ 
            \hline
            \verb|DeconvOptim.jl| & 14.0 & 0.86\\
            \verb|DeconvOptim.jl| - TV & 18.9 & 1.55 \\
            Richardson-Lucy (Julia) & 14.8 & 1.26 \\
            \verb|ThreeDeconv.jl|& 2415 & 44.5 \\
            \verb|GenericDeconv|  & 127 & 5.9\\
            \verb|GenericDeconv| - TV & 213 & 8.7 \\
            \verb|Huygens|  & & \\
            \verb|DeconvolutionLab2| - LR & 1089 & \\
            \verb|DeconvolutionLab2| - LR TV& 1131 & 
            \end{tabular}
            \caption{Runtimes of different deconvolution implementations.}
            \label{tab:deconv}
        \end{table}

        For the results of \verb|GenericDeconv| (written in Matlab by Rainer Heintzmann) we used a Poisson loss function with a L-BFGS minimizer and a quadratic non-negativity constraint.
        When the TV regularizer was added the weight was $\lambda = 0.002$.
        The general idea of \verb|GenericDeconv| is very similar to \verb|DeconvOptim.jl|.
        \verb|DeconvolutionLab2| offers a wide variety of different algorithms which however only run on a CPU. To put the numbers into context, even the non-iterative Tikhonov Regularization Inverse Filter (TRIF/TR) had a runtime of \SI{4}{\second}. In the table we show the results of an 
        iterative Richardson-Lucy deconvolution. As number of iterations we set 80 and as regularizer weight $\lambda=0.002$.\\
        In \verb|DeconvOptim.jl| we measured the runtime of the default deconvolution routine with a Poisson loss function
        and the runtime of straightforward Richardson-Lucy deconvolution without a regularizer.
        \verb|ThreeDeconv.jl| is another Julia deconvolution toolbox for deconvolution of three dimensional data.
        In contrast to their according publication \cite{ikoma2018convex} we needed more iterations to achieve good results
        on our used dataset. The regularizer weight was $\lambda=10^{-7}$. 
        The software requires a flat and a dark field image we could not provide. That might influence the quality of the reconstructed images
        but in terms of performance the runtimes might by only by a factor of 2-4 faster than the runtimes we obtained (Ikoma et al. used 140 iterations in their datasets).

        Since it is hard to judge the quality of those results with a numerical loss function value, we compare them visually.
        The CPU versions were multi threaded (if possible). \verb|DeconvolutionLab2|, \verb|Huygens| and \verb|GenericDeconv| were tested under Windows 10 and not Ubuntu.

        \begin{table}[h]
            \begin{tabular}{l l l}
                & Loss & iterations \\ 
            \hline
            \verb|DeconvOptim.jl| & Poisson & 15 \\
            \verb|DeconvOptim.jl| TV & Poisson & 15\\
            Richardson-Lucy & Poisson & 120 \\
            \verb|ThreeDeconv.jl|  & Gauss+Poisson& 500\\
            \verb|GenericDeconv|  & Poisson & 40\\
            \verb|GenericDeconv| TV & Poisson & 40 \\
            \verb|Huygens|  & & \\
            \verb|DeconvolutionLab2| LR & Poisson & 200\\
            \verb|DeconvolutionLab2| LR TV& Poisson & 200 
            \end{tabular}
            \caption{Overview which loss function (noise model) and number of iterations are used.}
            \label{tab:deconvprop}
        \end{table}

        As summary of those performance tests we can state that our Julia implementation is very competitive and especially on the CPU it is only
        beaten by the commercial \verb|Huygens| software. Despite \verb|ThreeDeconv.jl| written in Julia 
        it is notably slower in comparison to our implementation.
        \verb|DeconvolutionLab2| is written in Java and uses hand optimized gradients but still its performance is poor in comparison to all
        other packages.
        \begin{figure}[h]
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/DeconvOptim_cuda_no_regularizer_0.87s.png}
                \caption{\texttt{DeconvOptim.jl} without regularizer}
            \end{subfigure}
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/DeconvOptim_TV_0.002.png}
                \caption{\texttt{DeconvOptim.jl} with TV}
            \end{subfigure}
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/Lucy_richardson_julia.png}
                \caption{\texttt{DeconvOptim.jl} Richardson-Lucy}
            \end{subfigure}
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/Deconvolution_Lab2_200iterations.PNG}
                \caption{\texttt{Deconvolution\\Lab2} with TV}
            \end{subfigure}
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/ThreeDeconv_600_iterations.png}
                \caption{\texttt{ThreeDeconv.jl}}
            \end{subfigure}
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth, height=2.5cm]{example-image-a}
                \caption{\texttt{GenericDeconv}}
            \end{subfigure}
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth, height=2.5cm]{example-image-a}
                \caption{\texttt{Huygens}}
            \end{subfigure}
            \begin{subfigure}{.3\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/noisy.png}
                \caption{~~Measured image with Poisson noise}
            \end{subfigure}
            \caption{Overview of some of the deconvolution results of different packages.}
            \label{img:results}
        \end{figure}

\section{Summary}
    In conclusion, \verb|DeconvOptim.jl| is a very performant, flexible deconvolution software to deconvolve multi-dimensional datasets.
    Its focus has been from the beginning to be fast and the source code should be automatic differentiated by the available
    packages in Julia. Usually we would expect that AD decreases performances but still our software is faster than
    many other packages which have been written with hand-optimized gradients.
    Furthermore, our toolbox is not restricted in dimensionality and has mechanism to prevent wrap-around artifacts which
    occur with FFT based convolutions.
    Julia's metaprogramming capabilities together with \verb|Tullio.jl| allow to generate regularizer which have high performance
    and are easy to adapt.
    In future, we want to support more regularizers on GPUs since they can not be handled efficiently with the current \verb|Tullio.jl| solution.

\input{bib.tex}

\end{document}